---
title: "Are Aliens Real?"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
if (!requireNamespace("tidytuesdayR", quietly = TRUE)) {
  install.packages("tidytuesdayR")
}
library(tidytuesdayR)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

# Get the Data
# Reading the data manually

ufo_sightings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-20/ufo_sightings.csv')
places <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-20/places.csv')
day_parts_map <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-20/day_parts_map.csv')

```

## Introduction

**Description of the dataset**

The UFO Sightings Redux data comes from the National UFO Reporting Center, cleaned and enriched with data from sunrise-sunset.org by Jon Harmo.

The collection of verified UFO sightings from multiple sources is compiled into a comprehensive list called TidyTuesday UFO sightings. The location (latitude and longitude), the time and date of the sighting, the shape of the purported UFO, the duration of the sighting, and any other comments or descriptions provided by witnesses are typically included in this dataset.

There are 3 CSV files in the dataset:

-   `ufo_sightings.csv`: It has 12 variables such as city, state, country_code, and shape. In total 96,429 rows.

-   `places.csv`: It has 10 variables such as longitude, latitude and time zone, country_code, and alternate_city_names. There are 14,417 rows.

-   `day_parts_map.csv`: It has 12 variables such as sunrise, sunset, civil_twilight_end, nautical_twilight_end, and astronomical_twilight_end. It’s of 26,409 rows.

Make sure to load the data and use inline code for some of this information.

**Why did you choose this dataset?**

There are several reasons why we chose this dataset. The first reason is that the data is well-formatted and is rich in variables, making it interesting for data wrangling and visualizing patterns. The second reason is the social media engagement and interest in this particular phenomenon. We will determine whether the Naruto Run to Area 51 was worth the time and resources. Lastly, UFO sightings hold a plethora of significance in scientific research, as humans have been wondering and searching for life outside Earth ever since the first telescope was discovered. In summary, it would be interesting to engage in discussions of mysteries of the cosmos and extraterrestrial life while also contributing to the cause with our visualizations.

## Questions

Q.1) What is the geographic distribution of UFO sightings? Are there any hotspots where the number of sightings is prevalent? Is there any correlation between the location of the sighting and the shape of the UFO?

Q.2) Are there any patterns to the sightings concerning seasons? Do astronomical events affect the reported sightings?

## Analysis plan

**What is the geographic distribution of UFO sightings? Are there any hotspots where the number of sightings is prevalent? Is there any correlation between the location of the sighting and the shape of the UFO?**

-   Our plan for tackling this question involves using the `ufo_sightings.csv` and the `places.csv`.
-   The data will be then visualized, and the geographical locations of the sightings will be plotted on a map using the `ggplot2` package.
-   The hotspots where the sightings are prevalent can also be visualized using the dataset.
-   Then a stacked bar chart can be plotted to show the correlation between the location of the sightings and the shape of the UFO.

**Are there any patterns to the sightings concerning seasons? Do astronomical events affect the reported sightings?**
