---
title: "Proposal title"
subtitle: "Proposal"
format: html
editor: visual
---

```{r}
#| label: load-pkgs
#| message: false
library(tidyverse)
if (!requireNamespace("tidytuesdayR", quietly = TRUE)) {
  install.packages("tidytuesdayR")
}
library(tidytuesdayR)
```

## Dataset

```{r}
#| label: load-dataset
#| message: false

# Get the Data

# Read in with tidytuesdayR package 
# Install from CRAN via: install.packages("tidytuesdayR")
# This loads the readme and all the datasets for the week of interest

# Either ISO-8601 date or year/week works!

tuesdata <- tidytuesdayR::tt_load('2023-06-20')
tuesdata <- tidytuesdayR::tt_load(2023, week = 25)

ufo_sightings <- tuesdata$`ufo_sightings`
places <- tuesdata$`places`
day_parts_map <- tuesdata$`day_parts_map`

# Or read in the data manually

ufo_sightings <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-20/ufo_sightings.csv')
places <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-20/places.csv')
day_parts_map <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2023/2023-06-20/day_parts_map.csv')

```

A brief description of your dataset including its provenance, dimensions, etc. as well as the reason why you chose this dataset.

Make sure to load the data and use inline code for some of this information.

## Questions

Q.1) What is the geographic distribution of UFO sightings? Are there any hotspots where the number of sightings is prevalent? Is there any correlation between the location of the sighting and the shape of the UFO?

Q.2) Are there any patterns to the sightings concerning seasons? Do astronomical events affect the reported sightings?

## Analysis plan

-   A plan for answering each of the questions including the variables involved, variables to be created (if any), external data to be merged in (if any).
